---
title: "Function"
format: html
editor: visual
author: Sam Feinberg
---

# Exercise Functions:

## Question 1

In three or four sentences, explain why constructing a Function in order to execute tasks, is beneficial or advantageous.

Constructing a function to execute tasks is an extremely beneficial practice, as it makes the work of executing that task repeatable. By creating a function, a task can be repeated over and over, by simply plugging new inputs into the function. This way, you do not need to create a long new code for each iteration of the task, but can use the same code repeatedly.

## Question 2

Write a function (using r code and structure demonstrated in class) to calculate a z score for a given observed value, a mean, and a standard deviation value. And then use your function to find a z score for the following problem. (Research the internet to find the formula used to calculate a z score)\
Observed value = 25.77, mean = 23.54, standard deviation = 2.442

```{r}
#| title: Question 2

z_score <- function(observed_value, mean, standard_deviation) {
  (observed_value - mean) / standard_deviation
}

z_score(25.77, 23.54, 2.442)

```

## Question 3

Write a function (using r code and the structure demonstrated in class) to calculate \# the natural log of a number multiplied by the common log of the same number divided by the cube root of a given prime number.\
Use your function to find the answer if the number to be used for both log expressions is 32 and the given prime number is 11.\
Also use R code to round your answer to the nearest tenth

```{r}
#| title: Question 3

calc <- function(log_number, prime_number) {
  round((log(log_number) * log10(log_number)) / (prime_number ^ (1/3)), 2)
}

calc(32, 11)

```

## Question 4

Use and show R coding to calculate the standard deviation for each variable of the data table `mtcars` using the `Special For Loop Method`.

```{r}
#| title: Question 4

head(mtcars)

standard_deviation_vec <- vector(mode = "numeric", length = length(mtcars))
for(i in seq_along(mtcars)) {
  standard_deviation_vec[[i]] <- sd(mtcars[[i]], na.rm = TRUE)
  }
standard_deviation_vec

```
